Scheduling and Amortization for MPC

### 2022-02-19 paper_SIMD.tex
Cleaned up Basic Vectorization a bit.

\STATE \COMMENT {Phase 1: Raise dimension of scalar variables to corresponding loop nest. We may assume linear traversal of the MPC-source.}

\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i_1,...,i_k$}
\FOR {each $Y_i$}
\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y_i) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
\STATE \hspace{0.25cm} {same-level $\rightarrow$ add $Y'_i = Y_i$} \\
\STATE \hspace{0.25cm} {outer-to-inner $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{raise\_dim}(Y_i)$}
\STATE \hspace{0.25cm} {inner-to-outer $\rightarrow$ add $Y'_i[i_1,...,i_k] = \mathit{drop\_dim}(Y_i)$}
\ENDFOR

\COMMENT {Optimistically vectorize all. $\stackrel{\rightarrow}{i}$ means vectorized dimension.}
\STATE {change to $X[\stackrel{\rightarrow}{i_1},...\stackrel{\rightarrow}{i_k}] = \mathit{Op}(Y'_1[\stackrel{\rightarrow}{i_1},...,\stackrel{\rightarrow}{i_k}],Y'_2[\stackrel{\rightarrow}{i_1},...,\stackrel{\rightarrow}{i_k}])$ }
\ENDFOR

\COMMENT {Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up.}

@@

\WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
\STATE {merge $cl_1$ and $cl_2$}
\ENDWHILE
\FOR {each closure $cl$ (after merge)}
\STATE {create FOR $i_d = 0; ...$ loop}
\STATE {add $\phi$-nodes in $cl$ to header block}
\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
\STATE {add statements in $cl$ to loop body in some order of dependences}
\STATE \COMMENT {Dimension is not vectorizable: }
\STATE {change $\stackrel{\rightarrow}{i_d}$ to $i_d$ in all statements in loop}
\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
\ENDFOR
{
\color{red}
\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
\STATE {discard $\phi$-node if not used in any $cl$}
\ENDFOR
}
\ENDFOR

\COMMENT {Phase 3: Dimensionality reduction: remove unnecessary dimensionality. \ana{Better draft! Needs a bit more work.}}

\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing 
FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. } 
\FOR {each stmt and dimension $X[...i...] = ...$}
\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$} 
%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
\ENDFOR

\COMMENT {Now clean up drop\_dim and raise\_dim}
\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
\ENDFOR
\STATE {do (1) (extended) constant propagation, (2) copy propagation and (3) dead code elimination to get rid of redundant variables and raise and drop dimension statements}

\COMMENT {Phase 4:}
\STATE {add SIMD for simdified dimensions}

>>>>

\STATE \COMMENT { Phase 1: Raise dimension of scalar variables to corresponding loop nest. We can traverse stmts linearly in MPC-source. }

%\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
%\FOR {each $Y_n$}
\FOR {each def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$}
\STATE {case def-use edge $\mathit{stmt}' (\mbox{def of } Y) \rightarrow \mathit{stmt} (\mbox{def of } X)$ of} \\
\STATE \hspace{0.25cm} {same-level: $Y'$ is $Y$} \\
\STATE \hspace{0.25cm} {outer-to-inner: add $Y'[i,j,k] = \mathit{raise\_dim}(Y)$ at $\mathit{stmt'}$}
\STATE \hspace{0.25cm} {inner-to-outer: add $Y'[i,j,k] = \mathit{drop\_dim}(Y)$ at $\mathit{stmt}$}
\ENDFOR

\COMMENT { Optimistically vectorize all. $\stackrel{\rightarrow}{i}$ means vectorized dimension. }
\FOR {each MPC $\mathit{stmt}: X = \mathit{Op}(Y_1,Y_2)$ in loop $i,j,k$}
\STATE {change to $X[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j},\stackrel{\rightarrow}{k}] = \mathit{Op}(Y'_1[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j},\stackrel{\rightarrow}{k}],Y'_2[\stackrel{\rightarrow}{i},\stackrel{\rightarrow}{j},\stackrel{\rightarrow}{k}])$ }
\ENDFOR

\COMMENT { Phase 2: Recreating FOR loops for cycles; vectorizable statements hoisted up. }

@@

\WHILE {there are closure $cl_1$ and $cl_2$ that intersect}
\STATE {merge $cl_1$ and $cl_2$}
\ENDWHILE
\FOR {each closure $cl$ (after merge)}
\STATE {create FOR $i_d = 0; ...$ loop}
\STATE {add $\phi$-nodes in $cl$ to header block}
\STATE {{\color{red}{add target-less $\phi$-node for A if $cl$ updates array A}}}
\STATE {add statements in $cl$ to loop body in some order of dependences}
\STATE \COMMENT { Dimension is not vectorizable: }
\STATE {change $\stackrel{\rightarrow}{i_d}$ to $i_d$ in all statements in loop}
\STATE {treat FOR loop as monolith node: some def-use edges become same-level.}
\ENDFOR
{
\color{red}
\FOR {each target-less $\phi$-node $A_1 = \phi(A_0,A_k)$}
\STATE {in vectorizable stmts, replace use of $A_1$ with $A_0$}
\STATE {discard $\phi$-node if not used in any $cl$}
\ENDFOR
}
\ENDFOR

\COMMENT { Phase 3: Dimensionality reduction: removing unnecessary dimensionality. }

\COMMENT { A dimension $i$ is dead on exit from stmt $X[...i...] = ...$ if all def-uses with targets outside of the enclosing 
FOR $i=0 ...$ MOTION loop end at target (use) $X' = \mathit{drop\_dim}(X, i)$. } 
\FOR {each stmt and dimension $X[...i...] = ...$}
\STATE {if $i$ is a dead dimension on exit from stmt $X[...i...] = ...$, remove $i$ from $X$ (all defs and uses)} 
%\COMMENT{ A stmt v[...i] is a "dead dimension i" if all edges out of FOR i = 0; ... loop end at drop\_dim(v, i)}
\ENDFOR

\COMMENT { Now clean up drop\_dim and raise\_dim }
\FOR {each $X' =  \mathit{drop\_dim(X,i)}$ }
\STATE {replace with $X' = X$ if $i$ is dead in $X$.}
\ENDFOR
\STATE {do (1) (extended) constant propagation, (2) copy propagation and (3) dead code elimination to get rid of redundant variables and raise and drop dimension statements}

\COMMENT { Phase 4: }
\STATE {add SIMD for simdified dimensions}
